---
itle: "Peer Graded Assignment. Reproducible research"
Author: "Santiago Cárdenas Arciniegas"
Date: "12/7/2020"
Output: html_documentT
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **Peer Graded Assignment. Reproducible research**

**Author**: "Santiago Cárdenas Arciniegas"

**Date**: "12/7/2020"

# Part 1:
- What is mean total number of steps taken per day?


## Loading data
```{r}
data<-read.csv("activity.csv")

#Organizing data

Tsteps <- data.frame(tapply(data$steps, data$date, FUN=sum, na.rm=TRUE))
names(Tsteps)<-"steps"
```

## Creating Plot

```{r}

library(ggplot2)
g<-ggplot(data=Tsteps,aes(steps))

g+geom_histogram(fill="pink")+ 
  labs(title = "Total Steps per day", x = "Steps", y = "Frequency")

```

## Finding Mean and median

```{r}
mean(Tsteps$steps)
median(Tsteps$steps)
```

# Part 2
- What is the average daily activity pattern?

## Aggregating data
```{r}
avg_steps<-aggregate(data$steps,by=list(data$interval),
           FUN=mean,na.rm=TRUE)

names(avg_steps) <- c("intervals", "avgsteps")
```

## Creating plot
```{r}

g2<-ggplot(avg_steps,aes(intervals,avgsteps))

g2+geom_line(color="orange", size=1)+ labs(title = "Avg. Daily Steps", 
       x = "Interval", 
       y = "Avg. Steps per day")

```

- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

````{r}
avg_steps[which.max(avg_steps$avgsteps),]
````

# Part 3: 

- Calculate and report the total number of missing values in the dataset.

```{r}
length(which(is.na(data$steps)==TRUE))

```

-Devise a strategy for filling in all of the missing values in the dataset. 

## Creating new data set
```{r}
data2<-data
```

## Replacing NAS with For loop

```{r}
A<- nrow(data2)

data2$status <- "original"

for (i in 1:A) {
  if (is.na(data2[i,1])) {
  data2[i,1] <- mean(data$steps,na.rm = TRUE)
  data2[i,4] <- "completed"
    }
}
head(data2)

```

- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

## Arranging data by date

```{r}
Tsteps2 <- data.frame(tapply(data2$steps, 
                             data2$date, 
                             FUN=sum, 
                             na.rm=TRUE))

names(Tsteps2)<-"Steps2"
```

## Creating plot

```{r}

g3<-ggplot(Tsteps2,aes(Steps2))
g3+geom_histogram(fill="firebrick4")+
  labs(title = "Total number of steps taken each day\
       (NA replaced by mean)", x="Steps", y="Frecuency")

```

## Calculating mean and median
```{r}
mean(Tsteps2$Steps2)
median(Tsteps2$Steps2)
```

# Part 4: 

-Differences in activity patterns between weekdays and weekends

## Creating column in data set, that corresponds to the weekday

```{r}
library(gridExtra)
data2$date <- as.Date(data2$date)
data2$WD <- weekdays(data2$date)
A<- nrow(data2)

data2$Weekgroup<-"default"

head(data2)
```

## Adding "weekday" and "weeknd" entries in the Weekgroup column with for loop

```{r}
for (i in 1:A) {
  if (data2[i,5] == "sábado" | data2[i,5] == "domingo") {
    data2[i,6] <- "weekend"
  } else {data2[i,6]<-"weekday"}
}

head(data2)
```

-Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

## Agregating data by intervals and weekdays

```{r}
x_avg_steps<-aggregate(data2$steps,by=list(data2$interval,data2$Weekgroup),
                     FUN=mean,na.rm=TRUE)

names(x_avg_steps) <- c("intervals","weekgroup","avgsteps")

head(x_avg_steps)

```


## Creating plot

```{r}

g4<-ggplot(x_avg_steps,aes(intervals,avgsteps, color=`weekgroup`))


g4+geom_line(size=1)+ labs(title = "Avg. Daily Steps by Weekgroup", 
                                          x = "Interval", 
                                          y = "Avg. Steps per day")+
  facet_wrap(~x_avg_steps$weekgroup,ncol = 1, nrow=2)

```




